---
title             : "How infant sleep quality affects morning mood"
shorttitle        : "Infant sleep and mood"
author: 
  - name          : "Caspar Addyman"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, Goldsmiths, University of London, New Cross, London, SE14 6NW, UK"
    email         : "c.addyman@gold.ac.uk"
  - name          : "Frank Wiesemann"
    affiliation   : "2"
affiliation:
  - id            : "1"
    institution   : "Goldsmiths, University of London"
  - id            : "2"
    institution   : "R&D Baby Care, Procter & Gamble"
authornote: |
  Department of Psychology, Goldsmiths, University of London, New Cross, London, SE14 6NW, UK
abstract: |
  Infant sleep problems are among new parents greatest concerns. Even when sleep. 
  
  Sleep quantity and quality are recognised as important for infant development. But it remains an under-researched topic. Previous research has looked at the relationship between sleep and temperament and found that increased sleep length correlated with increased approachability and adaptability. 
  
  The present study uses sleep diaries with samples of parents and infants in Brazil and the United Kingdom to investigate what factors besides duration may influence sleep quality. 
  
  We look at how sleeping arrangements, night-time disturbances and diaper quality effected sleep duration and how these factors combined to influence morning mood. 
  Using linear models we find that in both samples infant bedtime is best predictor of sleep duration.  
  
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
keywords          : "keywords"
wordcount         : "X"
bibliography      : ["Sleep.bib", "r-references.bib"]
note              : Preprint submitted to peer-review on 1st May 2019. CRAN checkpoint date 2019-02-01
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
#first let's fix the snapshot date for all packages we will use as 1st March 2021
library(checkpoint)
checkpoint("2021-03-01")
# devtools::install_github("crsh/papaja")
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Infants sleep a great deal but with many wakings.\
Infants sleep affects parents sleep.
But little research investigates this.

Study of sixteen female subjects in a sleep lab for two weeks, found that subjective sense of good sleep was primarily related to sleep continuity @akerstedtSubjectiveMeaningGood1994

*For example, one study used the naturally occurring wakening of cloth diapers vs commercial diapers as a proxy measure of manipulated sleep disruption (Lukowski et al., 2015). Still,*

*Recent generations of British and American parents have been advised by pediatric health professionals that sleeping with their babies is "wrong" (e.g., Spock and Rothenberg 1985; Sturgess 1977; Wright 1972). (Ball et al., 1999)*

*Camerota et al. (2019) state that around 40% of reviewed infant sleep studies use only subjective, parent-report measures, such as a sleep diary or sleep questionnaires as indicators for infant sleep (Camerota et al., 2019). However, research has shown that parents are not very good at estimating their infants' sleep, often under- or overestimating- their infants' sleep compared to more objective measures (Camerota et al., 2018; Sadeh, 1994).*

*BISQ is good. = Good validity was established by Sadeh (2004) originally compared to actigraphy and sleep diaries.*

*Studies often ask parents if infants slept well. Here we are interested in whether the infants seemed happy in the morning. A more objective measure of infant emotional & behavioural state that might be useful correlate of sleep quality. Did Mindell use this?*

*Because*

*For instance, reactive infant temperament has been associated with shorter sleep\
duration and more night wakenings (Netsi et al., 2017) and an easy temperament (e.g.,\
increased approachability or low distractibility) was associated with longer overall\
objectively and parent-reported sleep duration (Kaley et al., 2012; Spruyt et al., 2008).\
Higher infant surgency (or an easy temperament) also moderated the impact of maternal\
emotional availability on sleep duration, suggesting a potential role of infant\
temperament on maternal mood and on sleep (Jian & Teti, 2016). Specifically, for infants\
with high surgency, higher maternal emotional availability was associated with higher\
sleep duration. No such effects were found for low surgency or the other subdimensions*

*For example, Mindell and colleagues (2010) report crosscultural differences in sleep duration and parent-reported sleep problems (Mindell et al.,\
2010). These differences in sleep duration are likely related to cross-cultural differences\
in how infants are put to sleep and availability of parents at bedtime and at night. Studies\
found that Brazilian infants and toddlers and children from Asian countries had\
systematically later bedtimes (by almost 2 hours) than were reported in European/US\
samples (Mindell et al., 2010; Netsi et al., 2017)*

# Methods

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

Middle class & lower middle class families recruited in southern central São Paulo.
Recruitment was done

116 mothers & babies (53 female, mean age = 13.9 months, range = 2-27m)

The study was conducted according to the guidelines in the Declaration of Helsinki and procedures were approved by the Ethics Committee of the Birkbeck, University of London.
Written informed consent was obtained from all parents or guardians prior to data collection.

## Material

On Day 0 all participants Family info Baby age Baby health screening Sleep arrangements What diaper brand?
Temperament (IBQ-R)

For 10 days participants kept a diary of infant sleep and morning mood (Supplementary Materials 1).

Amount of sleep.
How was diaper in morning?
Number of times woke up.
Feed?
Change?
Morning Happiness (Scale 1-10) Morning Energy (Scale 1-10)

## Procedure

## Data analysis

We used `r cite_r("r-references.bib")` for all our analyses.

# Results

The Rstats package glmulti was used to compare linear models predicting sleep duration from diary variables.
The best fitting linear model accounted for 27% of variance.
Bedtime, diaper absorbency, diaper change, diaper morning state were significant factors.

However, a simple model with just bedtime as a factor had R2 = 23%

Infant sleep problems are among new parents' greatest concerns and the importance of sleep quantity and quality for infant development is an under researched topic.
This project reports the results a survey of parents in São Paulo, Brazil.
The mothers of 117 infants (53 female, mean age = 13.9 months, range = 2-27m) provided background demographic data, general information on their child's sleep and completed the appropriate version of the short infant behaviour questionnaire (IBQ-R, Rothbart & Gartstein, 2000; EBQ, Putnam & Rothbart, 2006).
They also completed a 10 day sleep diary indicating the time babies went to sleep and woke up, night time wakes, feeds and diaper changes and the morning happiness and energy of their baby on a 10 point scale.

\
Preliminary analysis indicated that overall infants were in bed for an average of 9h46 ± 1h12 and woke up happy (mean score 8.2 +/-1.55) and energetic (mean score 7.2 +/-2.50).
A regression analysis showed that babies' morning energy level was positively affected by the number of night time wakings (beta=0.32, p\<.001) and total sleep (beta=0.42, p\<.001).
By contrast happiness was negatively affected by night time wakings (beta=-0.31, p\<.001) but showed an interaction between total sleep and diaper quality (total sleep: beta=0.13, p\<.003, interaction beta=-0.14, p\<.02).
These patterns are shown in Figure 1.
Sleep and morning mood were also affected by sleeping arrangements and infant temperament (not shown).
Overall, the data showed a complex relationship between infant sleep quality and morning mood but that parents can potentially improve morning mood by minimising night-time disturbances and using more absorbent diapers.

# Discussion

Infant nighttime sleep duration primarily predicted by bedtime.
Duration not affected by age, nighttime feeding or number of wakes and shows only small effects related to diaper.
Infant temperament does not appear to affect sleep or morning mood Babies morning energy increases with amount of sleep.
Happiness increases only in absorbent diapers

*A recent meta-analysis by Tomfohr-Madsen and colleagues (2020) revealed that the higher the neighbourhood SES of the child, the longer the objectively as well as subjectively reported sleep duration (Tomfohr-Madsen et al., 2020). For example, Matenchuk et al. (2019) found that 3-months old infants slept 40 minutes longer on average if the mother had a university degree (Matenchuk et al., 2019).*

# Conflicts of interest

Dr Addyman served as a paid consultant for Procter and Gamble.
Dr Wiesemann is an employee of Procter and Gamble

# Acknowledgements

The Brazillian study was supported by Ketchum PR, Developers Market Research.
The authors acknowledge the support of Cynitha OLivera The UK study was run by IPSOS Mori.

\newpage

# References

```{r create_r-references}
r_refs(file = "Sleep.bib")
```

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs}
:::

\endgroup
