---
title: "Baby Sleep Diary - Data Analysis"
author: "Caspar Addyman"
date: "26/04/2019"
output: html_document
---

This notebook sets up the `R` environment and packages to run analyses for sleep diary studies run in Sao Paulo, Brazil and the United Kingdom. Studies were conducted by Caspar Addyman [c.addyman@gold.ac.uk](mailto:c.addyman@gold.ac.uk) in collaboration with Procter & Gamble.

Note:
In this directory there are synchronised RMarkdown (`.Rmd`) and Jupyter/IPython (`.ipynb`) versions of this document. We create the Jupyter version by using the python package [`notedown`](https://github.com/aaren/notedown) with the following command:

```
notedown input.Rmd --knit > output.ipynb
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
```

## R Libraries

These are the libraries we use. We use `checkpoint` to fix all their versions to the MRAN snapshot for 1st May 2019.

The base version of R used was R-3.6.0.  

Note: The package `glmulti` depends on `rJava` which can be [quite fiddly to install]( 
https://cimentadaj.github.io/blog/2018-05-25-installing-rjava-on-windows-10/installing-rjava-on-windows-10/).



```{r libraries, include = TRUE}
library(checkpoint)
checkpoint("2019-05-01")
library(tidyverse)
library(foreign)
library(haven)
library(expss)
library(lubridate)
library(ggthemes)
library(corrgram)
library(gridExtra)
library(leaps)
library(MASS)
library(rJava)
library(glmulti)
```

